>

### 2019-04-09 paper/spinifex-paper.tex
remove variable backticks. fixed latex, summary section. grammerly,  knit.

\hypertarget{fix-here}{%
\subsection{FIX here}\label{fix-here}}

The steps are described in more detail below. The R functions used below
mentioned briefly, but more complete code examples are given with

@@

\begin{itemize}
  \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions.
  \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate.
  \item $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes.
  \item $\phi$, the angle of out-of-projection-plane rotation, coming into the manipulation space.
\end{itemize}

@@

In figure \ref{fig:step0} the contributions of the variables
\texttt{tars1} and \texttt{aede2} are mostly orthogonal to the
contributions of the other four variables. These two variables explain
the variation of the data distinguishing the purple group from the rest
of the sample. Select \texttt{aede2} as the manip var, the variable to
be manipulated, as it has a larger contribution to the projection. The
question that will be explored is how important the variable
\texttt{aede2} is to the separation of the clusters.

@@

Initialize a zero vector \textbf{e} of length \emph{p}. Set the fifth
element to one, as \texttt{aede2} is the fifth variable in the data,
giving the manip var a full contribution in this dimension. Use the
Gram-Schmidt process to orthonormalize the zero vector onto the basis
yielding the 3D manipulation space, \textbf{M}.

\begin{align*}
  \textbf{e} &\leftarrow Orthonormalize_{GS}(\textbf{e}) ~~~ w.r.t.~the~basis \\
  &\leftarrow \textbf{e} - \langle \textbf{e},\textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\
  \\
  \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e})
\end{align*}

@@

For a radial tour fix the (horizontal) angle within the projection
plane, \(\theta\), and define a sequence for the (vertical) angle coming
out of the projection plane, \(\phi\), bringing the initial \texttt{XY}
contributions of the manip var to a maximum and then to zero before
returning to the initial position. Dynamic capture of user manipulation
is typically performed directly on the projection plane (without
depiction of the manipulation space.)

@@

Figure \ref{fig:step3} illustrates a sequence with 15 projected bases,
showing the reference axes on top the corresponding projected data
points on the below. As a result, we can see that changes in the manip
var controlled the distance between the purple cluster and the remaining
sample, \texttt{aede2} is crucial in distinguishing this species. Tours
are typically viewed as an animation such a dynamic version of this tour
can be viewed online at
\url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}. The
page may take a moment to load.

@@

The data has a hierarchical structure with top-level clusters;
\texttt{DIS}, \texttt{VBP}, and \texttt{jet}. Each cluster is a
particular class of experiments, each with many experimental datasets
which, in turn, have many observations. In the consideration of data
density, we conduct manual tours on subsets of the \texttt{DIS} and
\texttt{jet} clusters. This explores the sensitivity of the structure to
each of the variables in turn and we present the subjectively best and
worst variable to manipulate for identifying dimensionality of the
clusters and describing the span of the clusters.

@@

The \texttt{jet} cluster resides in a smaller dimensionality than the
full set of experiments with four principal components explaining 95\%
of the variation in the cluster (Cook, Laa, and Valencia
\protect\hyperlink{ref-cook_dynamical_2018}{2018}). The data within this
4D embedding is subset down to \texttt{ATLAS7old} and \texttt{ATLAS7new}
to focus in on two groups with a reasonable number of observations that
occupy different parts of the subspace. Radial manual tours controlling
contributions from PC4 and PC3 are shown in figure
\ref{fig:JetClusterGood} and figure \ref{fig:JetClusterBad}
respectively. These components are selected to contrast the difference
of information conveyed by touring different variables. Links to dynamic
HTML5 animations controlling each of the four variables are also
provided.

When manipulating PC4, there is a clear difference in the parameter
space spanned by the experiment types \texttt{ATLAS7new} and
\texttt{ATLAS7old}. Yet, when PC3 is manipulated there is no indication
that the different experiments probe different parameter space.
Performing a radial manual tour on PC4 is more insightful than for PC3.

Radial manual tours manipulating each of the principal components in the
\texttt{jet} cluster can be viewed by following the links:
\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/}{PC1},
\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/}{PC2},
\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}{PC3},
and
\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/}{PC4}.

@@

(ref:JetClusterBad-cap) A radial manual tour of PC4 within the
\texttt{jet} cluster. Colored by experiment type: \texttt{ATLAS7new} in
green and \texttt{ATLAS7old} in orange. Data from \texttt{ATLAS7new}
(green) spans mostly the same space as \texttt{ALTLAS7old} (orange) with
no evident difference in cluster structure across varying contributions
of PC3. A dynamic version can be viewed at
\url{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}.

@@

A different space is used to explore the \texttt{DIS} cluster;
specifically the first six principal components, which explain 48\% of
the variation contained within the aggregated data(Cook, Laa, and
Valencia \protect\hyperlink{ref-cook_dynamical_2018}{2018}). Radial
manual tours are performed on PC6 and PC2 in figure
\ref{fig:DISclusterGood} and figure \ref{fig:DISclusterBad}
respectively.

@@

\texttt{DIS} cluster manual tours manipulating each of the principal
components can be viewed from the links:
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/}{PC1},
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/}{PC2},
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/}{PC3},
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/}{PC4},
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/}{PC5},
and
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/}{PC6}.

@@

\hypertarget{todo}{%
\subsection{TODO:}\label{todo}}

@@

@@

Tours, the dynamic linear projection of multivariate data, is an
important aspect of data visualization extending the display of
data-space as data dimensionality increases. This research has modified
the algorithm producing manual tours, applied this functionality in R
and offers extends the graphics offerings that can be used to display
tours. The paragraphs below explore how this work might be extended.

>>>>

The steps are described in more detail below with description,
equations, figures, and their code usage.

@@

\begin{itemize}
  \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions.
  \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate, the manip var.
  \item $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes.
  \item $\phi$, the angle of out-of-projection-plane rotation, coming into the manipulation space.
\end{itemize}

@@

In figure \ref{fig:step0} the contributions of the variables tars1 and
aede2 are mostly orthogonal to the contributions of the other four
variables. These two variables explain the variation of the data
distinguishing the purple group from the rest of the sample. Select
aede2 as the manip var, the variable to be manipulated, as it has a
larger contribution to the projection. The question that will be
explored is how important the variable aede2 is to the separation of the
clusters.

@@

Initialize a zero vector \textbf{e} of length \emph{p}. Set the fifth
element to one, as aede2 is the fifth variable in the data, giving the
manip var a full contribution in this dimension. Use the Gram-Schmidt
process to orthonormalize the zero vector onto the basis yielding the 3D
manipulation space, \textbf{M}.

\begin{align*}
  \textbf{e}_k &\leftarrow 1 \\ 
  \textbf{e}   &\leftarrow \textbf{e} - \langle \textbf{e}, \textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\ 
  &\\
  \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e})
\end{align*}

@@

For a radial tour fix the (horizontal) angle within the projection
plane, \(\theta\), and define a sequence for the (vertical) angle coming
out of the projection plane, \(\phi\), bringing the initial \(XY\)
contributions of the manip var to a maximum and then to zero before
returning to the initial position. Dynamic capture of user manipulation
is typically performed directly on the projection plane (without
depiction of the manipulation space.)

@@

Figure \ref{fig:step3} illustrates a sequence with 15 projected bases,
showing the reference axes on top the corresponding projected data
points on the below. As a result, we can see that changes in the manip
var controlled the distance between the purple cluster and the remaining
sample, aede2 is crucial in distinguishing this species. Tours are
typically viewed as an animation such a dynamic version of this tour can
be viewed online at
\url{https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/}. The
page may take a moment to load.

@@

The data has a hierarchical structure with top-level clusters; DIS, VBP,
and jet. Each cluster is a particular class of experiments, each with
many experimental datasets which, in turn, have many observations. In
the consideration of data density, we conduct manual tours on subsets of
the DIS and jet clusters. This explores the sensitivity of the structure
to each of the variables in turn and we present the subjectively best
and worst variable to manipulate for identifying dimensionality of the
clusters and describing the span of the clusters.

@@

The jet cluster resides in a smaller dimensionality than the full set of
experiments with four principal components explaining 95\% of the
variation in the cluster (Cook, Laa, and Valencia
\protect\hyperlink{ref-cook_dynamical_2018}{2018}). The data within this
4D embedding is subset down to ATLAS7old and ATLAS7new to focus in on
two groups with a reasonable number of observations that occupy
different parts of the subspace. Radial manual tours controlling
contributions from PC4 and PC3 are shown in figure
\ref{fig:JetClusterGood} and figure \ref{fig:JetClusterBad}
respectively. These components are selected to contrast the difference
of information conveyed by touring different variables. Links to dynamic
HTML5 animations controlling each of the four variables are also
provided.

When manipulating PC4, there is a clear difference in the parameter
space spanned by the experiment types ATLAS7new and ATLAS7old. Yet, when
PC3 is manipulated there is no indication that the different experiments
probe different parameter space. Performing a radial manual tour on PC4
is more insightful than for PC3.

Radial manual tours manipulating each of the principal components in the
jet cluster can be viewed by following the links:
\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/}{PC1},
\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/}{PC2},
\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}{PC3},
and
\href{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/}{PC4}.

@@

(ref:JetClusterBad-cap) A radial manual tour of PC4 within the jet
cluster. Colored by experiment type: ATLAS7new in green and ATLAS7old in
orange. Data from ATLAS7new (green) spans mostly the same space as
ALTLAS7old (orange) with no evident difference in cluster structure
across varying contributions of PC3. A dynamic version can be viewed at
\url{https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/}.

@@

A different space is used to explore the DIS cluster; specifically the
first six principal components, which explain 48\% of the variation
contained within the aggregated data(Cook, Laa, and Valencia
\protect\hyperlink{ref-cook_dynamical_2018}{2018}). Radial manual tours
are performed on PC6 and PC2 in figure \ref{fig:DISclusterGood} and
figure \ref{fig:DISclusterBad} respectively.

@@

DIS cluster manual tours manipulating each of the principal components
can be viewed from the links:
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/}{PC1},
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/}{PC2},
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/}{PC3},
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/}{PC4},
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/}{PC5},
and
\href{https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/}{PC6}.

@@

@@

Tours, the dynamic linear projection of multivariate data, play an
important role in data visualization; they extending the dimensionality
of visuals while data- and parameter-spaces become ever larger. This
research has modified the algorithm producing manual tours and has made
this functionality available in package \textbf{spinifex}. The package
adds to \textbf{tourr}, extending the graphics offerings that can be
used to display tours.

Radial manual tours were applied to a dataset across different
experiments of hadronic collisions. The importance of selecting the
correct variable to manipulate is demonstrated by comparing tours of
varying quality. The manual tours convey a better picture of the
structure of the clustering than static linear or non-linear projections
are able to. Giving the full contribution of the manipulation space to
the manip var enables analysts to explore the sensitivity of the
structure with the selected variable. This information can be used by
physicists to which experiments are probing which parameter spaces,
which can be indispensable for smarter planning of effort and funding.

@@

### 2019-04-09 paper/spinifex_paper.Rmd
remove variable backticks. fixed latex, summary section. grammerly,  knit.

## FIX here
The steps are described in more detail below. The R functions used below mentioned briefly, but more complete code examples are given with

@@

<!-- remember mixing latex and rmarkdown bullets isn't trivial. -->
\begin{itemize}
  \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions.
  \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate.
  \item $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes.
  \item $\phi$, the angle of out-of-projection-plane rotation, coming into the manipulation space.
\end{itemize}

@@

<!-- select a manip var-->
In figure \ref{fig:step0} the contributions of the variables `tars1` and `aede2` are mostly orthogonal to the contributions of the other four variables. These two variables explain the variation of the data distinguishing the purple group from the rest of the sample. Select `aede2` as the manip var, the variable to be manipulated, as it has a larger contribution to the projection. The question that will be explored is how important the variable `aede2` is to the separation of the clusters.

@@

<!-- Zero Vect, manip sp -->
Initialize a zero vector **e** of length *p*. Set the fifth element to one, as  `aede2` is the fifth variable in the data, giving the manip var a full contribution in this dimension. Use the Gram-Schmidt process to orthonormalize the zero vector onto the basis yielding the 3D manipulation space, **M**.

\begin{align*}
  \textbf{e_k} &\leftarrow 1 \\
  \textbf{e}   &\leftarrow \textbf{e} - \langle \textbf{e},\textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\
  \\
  \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e})
\end{align*}

@@

<!-- describe manip var path -->
For a radial tour fix the (horizontal) angle within the projection plane, $\theta$, and define a sequence for the (vertical) angle coming out of the projection plane, $\phi$, bringing the initial `XY` contributions of the manip var to a maximum and then to zero before returning to the initial position. Dynamic capture of user manipulation is typically performed directly on the projection plane (without depiction of the manipulation space.)

@@

<!-- description of slides, on the technical side we haven't gone back to data space yet. -->
Figure \ref{fig:step3} illustrates a sequence with 15 projected bases, showing the reference axes on top the corresponding projected data points on the below. As a result, we can see that changes in the manip var controlled the distance between the purple cluster and the remaining sample, `aede2` is crucial in distinguishing this species. Tours are typically viewed as an animation such a dynamic version of this tour can be viewed online at https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/. The page may take a moment to load.

@@

<!-- Data structure -->
The data has a hierarchical structure with top-level clusters; `DIS`, `VBP`, and `jet`. Each cluster is a particular class of experiments, each with many experimental datasets which, in turn, have many observations. In the consideration of data density, we conduct manual tours on subsets of the `DIS` and `jet` clusters. This explores the sensitivity of the structure to each of the variables in turn and we present the subjectively best and worst variable to manipulate for identifying dimensionality of the clusters and describing the span of the clusters.

@@

<!-- jet cluster, explain dimensionality  -->
The `jet` cluster resides in a smaller dimensionality than the full set of experiments with four principal components explaining 95% of the variation in the cluster [@cook_dynamical_2018]. The data within this 4D embedding is subset down to `ATLAS7old` and `ATLAS7new` to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace. Radial manual tours controlling contributions from PC4 and PC3 are shown in figure \ref{fig:JetClusterGood} and figure \ref{fig:JetClusterBad} respectively. These components are selected to contrast the difference of information conveyed by touring different variables. Links to dynamic HTML5 animations controlling each of the four variables are also provided.

<!-- discussion of findings and which is more insightful -->
When manipulating PC4, there is a clear difference in the parameter space spanned by the experiment types `ATLAS7new` and `ATLAS7old`. Yet, when PC3 is manipulated there is no indication that the different experiments probe different parameter space. Performing a radial manual tour on PC4 is more insightful than for PC3.

Radial manual tours manipulating each of the principal components in the `jet` cluster can be viewed by following the links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).

@@

(ref:JetClusterBad-cap) A radial manual tour of PC4 within the `jet` cluster. Colored by experiment type: `ATLAS7new` in green and `ATLAS7old` in orange. Data from `ATLAS7new` (green) spans mostly the same space as `ALTLAS7old` (orange) with no evident difference in cluster structure across varying contributions of PC3. A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/.

@@

<!-- introduce DIS cluster -->
A different space is used to explore the `DIS` cluster; specifically the first six principal components, which explain 48% of the variation contained within the aggregated data[@cook_dynamical_2018]. Radial manual tours are performed on PC6 and PC2 in figure \ref{fig:DISclusterGood} and figure \ref{fig:DISclusterBad} respectively.

@@

`DIS` cluster manual tours manipulating each of the principal components can be viewed from the links: [PC1](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/), [PC4](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/), [PC5](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/), and [PC6](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/).

@@

## TODO:
### Summary

@@

<!-- summary and segue  -->
Tours, the dynamic linear projection of multivariate data, is an important aspect of data visualization extending the display of data-space as data dimensionality increases. This research has modified the algorithm producing manual tours, applied this functionality in R and offers extends the graphics offerings that can be used to display tours. The paragraphs below explore how this work might be extended.

>>>>

The steps are described in more detail below with description, equations, figures, and their code usage.

@@

<!-- remember mixing latex and rmarkdown bullets isn't trivial. -->
\begin{itemize}
  \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions.
  \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate, the manip var.
  \item $\theta$, the angle of in-projection-plane rotation, for example, on the reference axes.
  \item $\phi$, the angle of out-of-projection-plane rotation, coming into the manipulation space.
\end{itemize}

@@

<!-- select a manip var-->
In figure \ref{fig:step0} the contributions of the variables tars1 and aede2 are mostly orthogonal to the contributions of the other four variables. These two variables explain the variation of the data distinguishing the purple group from the rest of the sample. Select aede2 as the manip var, the variable to be manipulated, as it has a larger contribution to the projection. The question that will be explored is how important the variable aede2 is to the separation of the clusters.

@@

<!-- Zero Vect, manip sp -->
Initialize a zero vector **e** of length *p*. Set the fifth element to one, as  aede2 is the fifth variable in the data, giving the manip var a full contribution in this dimension. Use the Gram-Schmidt process to orthonormalize the zero vector onto the basis yielding the 3D manipulation space, **M**.

\begin{align*}
  \textbf{e}_k &\leftarrow 1 \\ 
  \textbf{e}   &\leftarrow \textbf{e} - \langle \textbf{e}, \textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\ 
  &\\
  \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e})
\end{align*}

@@

<!-- describe manip var path -->
For a radial tour fix the (horizontal) angle within the projection plane, $\theta$, and define a sequence for the (vertical) angle coming out of the projection plane, $\phi$, bringing the initial $XY$ contributions of the manip var to a maximum and then to zero before returning to the initial position. Dynamic capture of user manipulation is typically performed directly on the projection plane (without depiction of the manipulation space.)

@@

<!-- description of slides, on the technical side we haven't gone back to data space yet. -->
Figure \ref{fig:step3} illustrates a sequence with 15 projected bases, showing the reference axes on top the corresponding projected data points on the below. As a result, we can see that changes in the manip var controlled the distance between the purple cluster and the remaining sample, aede2 is crucial in distinguishing this species. Tours are typically viewed as an animation such a dynamic version of this tour can be viewed online at https://nspyrison.netlify.com/thesis/flea_manualtour_mvar5/. The page may take a moment to load.

@@

<!-- Data structure -->
The data has a hierarchical structure with top-level clusters; DIS, VBP, and jet. Each cluster is a particular class of experiments, each with many experimental datasets which, in turn, have many observations. In the consideration of data density, we conduct manual tours on subsets of the DIS and jet clusters. This explores the sensitivity of the structure to each of the variables in turn and we present the subjectively best and worst variable to manipulate for identifying dimensionality of the clusters and describing the span of the clusters.

@@

<!-- jet cluster, explain dimensionality  -->
The jet cluster resides in a smaller dimensionality than the full set of experiments with four principal components explaining 95% of the variation in the cluster [@cook_dynamical_2018]. The data within this 4D embedding is subset down to ATLAS7old and ATLAS7new to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace. Radial manual tours controlling contributions from PC4 and PC3 are shown in figure \ref{fig:JetClusterGood} and figure \ref{fig:JetClusterBad} respectively. These components are selected to contrast the difference of information conveyed by touring different variables. Links to dynamic HTML5 animations controlling each of the four variables are also provided.

<!-- discussion of findings and which is more insightful -->
When manipulating PC4, there is a clear difference in the parameter space spanned by the experiment types ATLAS7new and ATLAS7old. Yet, when PC3 is manipulated there is no indication that the different experiments probe different parameter space. Performing a radial manual tour on PC4 is more insightful than for PC3.

Radial manual tours manipulating each of the principal components in the jet cluster can be viewed by following the links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).

@@

(ref:JetClusterBad-cap) A radial manual tour of PC4 within the jet cluster. Colored by experiment type: ATLAS7new in green and ATLAS7old in orange. Data from ATLAS7new (green) spans mostly the same space as ALTLAS7old (orange) with no evident difference in cluster structure across varying contributions of PC3. A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/.

@@

<!-- introduce DIS cluster -->
A different space is used to explore the DIS cluster; specifically the first six principal components, which explain 48% of the variation contained within the aggregated data[@cook_dynamical_2018]. Radial manual tours are performed on PC6 and PC2 in figure \ref{fig:DISclusterGood} and figure \ref{fig:DISclusterBad} respectively.

@@

DIS cluster manual tours manipulating each of the principal components can be viewed from the links: [PC1](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc3/), [PC4](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc4/), [PC5](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc5/), and [PC6](https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/).

@@

### Summary

<!-- summary and segue  -->
Tours, the dynamic linear projection of multivariate data, play an important role in data visualization; they extending the dimensionality of visuals while data- and parameter-spaces become ever larger. This research has modified the algorithm producing manual tours and has made this functionality available in package **spinifex**. The package adds to **tourr**, extending the graphics offerings that can be used to display tours.

Radial manual tours were applied to a dataset across different experiments of hadronic collisions. The importance of selecting the correct variable to manipulate is demonstrated by comparing tours of varying quality. The manual tours convey a better picture of the structure of the clustering than static linear or non-linear projections are able to. Giving the full contribution of the manipulation space to the manip var enables analysts to explore the sensitivity of the structure with the selected variable. This information can be used by physicists to which experiments are probing which parameter spaces, which can be indispensable for smarter planning of effort and funding.

@@
